
class: inverse, middle

# Week 9: .fancy[Intro to Python] `r fontawesome::fa("python", fill = "white")`

## 1. Getting started
## 2. .orange[Python basics]
## 3. Functions & methods
## 4. Loops & lists
## 5. Strings

---
# Operators

```{css, echo=F}
    .remark-slide table{
        width: 100%;
    }
```

--
.leftcol[.left[
## Arithmetic operators

Operator | R      | Python
-----------------|---------|-----------
Integer division |  `%/%` | `//`
Modulus          |  `%%`  | `%`
Powers           |  `^`   | `**`
]]
--
.rightcol[.left[
## Logical operators

Operator  | R         | Python
----------|-----------|-----------
And       |  `&`      | `and`; `&`
Or        |  &#124;   | `or`; &#124;
Not       |  `!`      | `not`; `!`

You can do this in Python:
```{r, eval=FALSE}
(3 == 3) and (4 == 4)
```
]]

---
## Variable assignment

Python only uses the `=` symbol to make assignments (no `<-`):

```{python}
value = 3
value
```

---
## Data types

Same data types as R, but with more "Computer Science-y" names:

.leftcol60[.left[
Description          | R            | Python
---------------------|--------------|-----------
numeric (w/decimal)  | `double`   | `float`
integer              | `integer`  | `int`
character            | `character`| `str`
logical              | `logical`  | `bool`
]]


---
## Data types

Three important distinctions:

.left[
Data type     | R                  | Python
--------------|--------------------|-----------
Logical       | `TRUE` or `FALSE` | `True` or `False`
Numbers       | `double` by default  | `int` by default (unless has decimal)
Nothing       | `NULL`| `None`
]

---
.leftcol[
.center[**R**: Get type with `typeof()`]
```{r}
typeof(3.14)
```
```{r}
typeof(3L)
```
```{r}
typeof("3")
```
```{r}
typeof(TRUE)
```
]
.rightcol[
.center[**Python**: Get type with `type()`]
```{python}
type(3.14)
```
```{python}
type(3)
```
```{python}
type("3")
```
```{python}
type(True)
```
]

---
.leftcol[
.center[**R**: _Check_ type with `is.type()`]
```{r}
is.double(3.14)
```
```{r}
is.integer(3L)
```
```{r}
is.character("3")
```
```{r}
is.logical(TRUE)
```
]
.rightcol[
.center[**Python**: _Check_ type with `type() == type`]
```{python}
type(3.14) == float
```
```{python}
type(3) == int
```
```{python}
type("3") == str
```
```{python}
type(True) == bool
```
]

---
.leftcol[
.center[**R**: _Convert_ type with `as.type()`]
```{r}
as.double("3")
```
```{r}
as.integer(3.14)
```
```{r}
as.character(3.14)
```
```{r}
as.logical(3.14)
```
]
.rightcol[
.center[**Python**: _Convert_ type with `type()`]
```{python}
float("3")
```
```{python}
int(3.14)
```
```{python}
str(3.14)
```
```{python}
bool(3.14)
```
]

---
## Quick practice

```{r, echo=FALSE}
countdown(minutes = 2,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

Write Python code to do the following:

1. Create an object `x` that stores the value `"123"`
2. Create an object `y` that is `x` converted to an integer
3. Write code to confirm that `y` is indeed an integer
4. Write a logical statement to determine if `y` is odd or even
