
class: inverse, middle

# Week 9: .fancy[Intro to Python] `r fontawesome::fa("python", fill = "white")`

## 1. Getting started
## 2. Python basics
## 3. .orange[Functions & methods]
## 4. Loops & lists
## 5. Strings

---
# Python and R have many similar functions

--
.leftcol[
.center[**R**]
```{r}
abs(-1)
round(3.14)
round(3.14, 1)
```
]
.rightcol[
.center[**Python**]
```{python}
abs(-1)
round(3.14)
round(3.14, 1)
```
]

---
# Writing functions

.leftcol[
.center[**R**]
```{r}
isEven <- function(n) {
    if (n %% 2 == 0) {
        return(TRUE)
    }
    return(FALSE)
}
```
]
--
.rightcol[
.center[**Python**]
```{python}
def isEven(n):
    if (n % 2 == 0):
        return(True)
    return(False)
```
Note:

- Functions start with `def`
- Use `:` and indentation instead of `{}`
- **Indentation is precisely 4 spaces!**
]

---
# Writing test functions

.leftcol[
.center[**R**]
```{r}
test_isEven <- function() {
    cat("Testing isEven(n)...")
    stopifnot(isEven(2) == TRUE)
    stopifnot(isEven(1) == FALSE)
    cat("Passed!")
}
```
]
--
.rightcol[
.center[**Python**]
```{python}
def test_isEven():
    print("Testing isEven(n)...")
    assert(isEven(2) == True)
    assert(isEven(1) == False)
    print("Passed!")
```
Note:

- Use `print()` instead of `cat()`
- Use `assert()` instead of `stopifnot()`
]

---
# Python Methods

Python objects have "methods" - special functions that _belong_ to certain object classes.

--
Example: Make a string upper case

.leftcol[
.center[**R**]
Use `str_to_upper()` function
```{r}
s <- "foo"
stringr::str_to_upper(s)
```
]
--
.rightcol[
.center[**Python**]
Use `upper()` _method_
```{python}
s = "foo"
s.upper()
```
]

---
# Python Methods

See all the available methods with `dir` function:
```{python}
s = "foo"
dir(s)
```

---
# R-Python magic

You can source a Python script from R, then use the Python function in R!


--
Inside your `notes.py` file, you have the following function defined:
```{python}
def isEven(n):
    if (n % 2 == 0):
        return(True)
    return(False)
```

--
Open your `notes.R` file and _source_ the `note.py` file:
```{r, eval=FALSE}
reticulate::source_python('notes.py')
```

--
Magically, the function `isEven(n)` now works inside R!

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

# Think-Pair-Share

Write the following two functions in Python code:

1. `hypotenuse(a, b)`: Returns the hypotenuse of the two lines of length `a` and `b`.

2. `isRightTriangle(a, b, c)`: Returns `True` if the triangle formed by the lines of length `a`, `b`, and `c` is a right triangle and `False` otherwise. **Hint**: you may not know which value (`a`, `b`, or `c`) is the hypotenuse.

```{python, echo=FALSE}
def hypotenuse(a, b):
    return((a**2 + b**2)**0.5)

def isRightTriangle(a, b, c):
    test1 = (c == hypotenuse(a, b))
    test2 = (b == hypotenuse(a, c))
    test3 = (a == hypotenuse(b, c))
    return(test1 or test2 or test3)
```
