
class: inverse, middle

# Week 9: .fancy[Intro to Python] `r fontawesome::fa("python", fill = "white")`

## 1. Getting started
## 2. Python basics
## 3. Functions & methods
## 4. .orange[Loops & lists]
## 5. Strings

---
## `for` loops

.leftcol[
.center[**R**]
```{r}
for (i in seq(1, 5, 2)) {
    cat(i, '\n')
}
```
]
--
.rightcol[
.center[**Python**]
```{python}
for i in range(1, 5, 2):
    print(i)
```
Notes:
- `range()` leaves out stopping number
- No `()` in `for` loop line
]

---
## `while` loops

.leftcol[
.center[**R**]
```{r}
i <- 1
while (i <= 5) {
    print(i)
    i <- i + 2
}
```
]
--
.rightcol[
.center[**Python**]
```{python}
i = 1
while i <= 5:
    print(i)
    i += 2
```
Notes:

- Could also use `i = i + 2` to increment
- No `()` in `while` loop line
]

---
# Python lists

**Note**: These are not the same as R vectors! (They're equivalent to R lists)


--
Universal list creator: `[]`
```{python}
[1, 2, 3]
```

--
Lists can store different types
```{python}
[1, "foo", True]
```

---
## Adding and removing items

--
.leftcol[
Add items with `list.append()`
```{python}
x = [1, 2, 3]
x.append(7)
x
```
**Note**: You don't have to overright `a`,<br>
i.e. Don't do this: `x = x.append(7)`
]
--
.rightcol[
Remove items with `list.remove()`
```{python}
x = [1, 2, 3]
x.remove(3)
x
```
]

---
## Sorting lists

```{python}
x = [1, 5, 3]
```
--
.leftcol[
Sorting that returns a new object
```{python}
sorted(x)
sorted(x, reverse = True)
x
```
]
--
.rightcol[
Sort the object `x` _without_ creating a new object
```{python}
x.sort()
x
```
]

---
# Slicing lists with `[]`

```{python}
x = ['A', 'list', 'of', 'words']
```
--
.leftcol[.code80[
Indices start at 0:
```{python}
x[0] # Returns the first element
```
```{python}
x[3] # Returns the third element
```
```{python}
x[len(x)-1] # Returns the last element
```
]]
--
.rightcol[.code80[
Slicing with a vector of indices:
```{python}
x[0:3]  # Returns the first 3 elements
```
]]

---
# Negative indices slice from the end

```{python}
x = ['A', 'list', 'of', 'words']
```
--
.leftcol[.code80[
Indices start at 0:
```{python}
x[-1] # Returns the last element
```
```{python}
x[-2] # Returns 2nd-to-last element
```
```{python}
x[-len(x)] # Returns first element
```
]]
--
.rightcol[.code80[
Slicing with a vector of indices:
```{python}
x[-3:-1]  # Returns middle 2 elements
```
]]

---
## Note on 0 indexing

```{python}
x = ["A", "B", "C", "D", "E"]
```
--
List items sit _between_ fence posts.
```{python, eval=FALSE}
index: 0     1     2     3     4
       |     |     |     |     |     |
item:  | "A" | "B" | "C" | "D" | "E" |
       |     |     |     |     |     |
```

--
You slice at the _fence post_ number to get elements _between_ the posts.
.leftcol[
```{python}
x[0:1]
```
]
.rightcol[
```{python}
x[0:3]
```
]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

# Think-Pair-Share

Write the following two functions in Python code:

1. `factorial(n)`: Returns the factorial of `n`, e.g. `3! = 3*2*1 = 6`. Note that `0` is a special case, and `0! = 1`. Assume `n >= 0`.

2. `nthHighestValue(n, x)`: Returns the nth highest value in a list of numbers. For example, if `x = [5, 1, 3]`, then `nthHighestValue(1, x)` should return `5`, because `5` is the 1st highest value in `x`, and `nthHighestValue(2, x)` should return `3` because it's the 2nd highest value in `x`. Assume that `n <= len(x)`.

```{python, echo=FALSE}
def factorial(n):
    if (n == 0):
        return(1)
    result = 1
    for i in range(n-1):
        result = result*i
    return(result)

def nthHighestValue(n, x):
    y = sorted(x, reverse = True)
    return(y[n - 1])
```
