
class: inverse, middle

# Week 9: .fancy[Intro to Python] `r fontawesome::fa("python", fill = "white")`

## 1. Getting started
## 2. Python basics
## 3. Functions & methods
## 4. Loops & lists
## 5. .orange[Strings]

---
## Doing "math" with strings

--
.leftcol[
Concatenation:
.center[**R**]
```{r}
paste("foo", "bar", sep = "")
```
]
--
.rightcol[
<br>
.center[**Python**]
```{python}
"foo" + "bar"
```
]
--
.leftcol[
Repetition:
.center[**R**]
```{r}
str_dup("foo", 3)
```
]
--
.rightcol[
<br>
.center[**Python**]
```{python}
"foo" * 3
```
]

---
## Using English with strings

.leftcol[
Sub-string detection:
.center[**R**]
```{r}
str_detect('Apple', 'App')
```
]
--
.rightcol[
<br>
.center[**Python**]
```{python}
'App' in 'Apple'
```
]

---
class: inverse, middle, center

## Most string manipulation is done with _methods_

--
.leftcol[
.center[**R**]
```{r, eval=FALSE}
str_function(s)
```
]
--
.rightcol[
.center[**Python**]
```{python, eval=FALSE}
s.method()
```
]

---
## Case conversion

--
.leftcol[
.center[**R**]
```{r}
s <- "A longer string"
str_to_upper(s)
str_to_lower(s)
str_to_title(s)
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = "A longer string"
s.upper()
s.lower()
s.title()
```
]

---
## Trimming white space

--
.leftcol[
.center[**R**]
```{r}
s <- "     A string with space     "
str_trim(s)
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = "     A string with space     "
s.strip()
```
]

---
## Replacing strings

--
.leftcol[
.center[**R**]
```{r}
s <- "Hello world"
str_replace(s, "o", "a")
str_replace_all(s, "o", "a")
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = "Hello world"
s.replace("o", "a")
```
]

---
## Merge a vector / list of strings together

--
.leftcol[
.center[**R**]
```{r}
s <- c("Hello", "world")
paste(s, collapse = "")
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = ["Hello", "world"]
"".join(s)
```
]

---
## Python has some super handy string methods

--
Detect if string contains only numbers:
.leftcol[
.center[**R**]
R doesn't have a function for this...<br>here's one way to do it:
```{r}
s <- "42"
! is.na(as.numeric(s))
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = "42"
s.isnumeric()
```
]

---
## Getting sub-strings with `[]`

--
.leftcol[
.center[**R**]
```{r}
s <- "Apple"
str_sub(s, 1, 3)
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = "Apple"
s[0:3]
```
Notes:

- Indexing is the same as lists
]

---
## Getting sub-string indices

--
.leftcol[
.center[**R**]
```{r}
s <- "Apple"
str_locate(s, "pp")
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = "Apple"
s.index("pp")
```
Note:

- Only returns the starting index
]

---
## String splitting

Both languages return a list:

--
.leftcol[
.center[**R**]
```{r}
s <- "Apple"
str_split(s, "pp")
```
]
--
.rightcol[
.center[**Python**]
```{python}
s = "Apple"
s.split("pp")
```
]

---
## Python can only split individual strings

--
.leftcol[
R can split vectors of strings
```{r}
s <- c("Apple", "Snapple")
str_split(s, "pp")
```
]
--
.rightcol[
.center[**Python**]
```{python, error=TRUE}
s = ["Apple", "Snapple"]
s.split("pp")
```
]

---
## Need **numpy** package for this in Python

```{python}
import numpy as np

s = np.array(["Apple", "Snapple"])
np.char.split(s, "pp")
```

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

# Think-Pair-Share

.font80[
Write the following two functions in Python code:

1. `sortString(s)`: Takes a string `s` and returns back an alphabetically sorted string. **Hint**: Use `list(s)` to break a string into a list of letters.

- `sortString("cba") == "abc"`
- `sortString("abedhg") == "abdegh"`
- `sortString("AbacBc") == "ABabcc"`

2. `areAnagrams(s1, s2)`: Takes two strings, `s1` and `s2`, and returns `True` if the strings are [anagrams](https://en.wikipedia.org/wiki/Anagram), and `False` otherwise. **Treat lower and upper case as the same letters**.

- `areAnagrams("", "") == True`
- `areAnagrams("aabbccdd", "bbccddee") == False`
- `areAnagrams("TomMarvoloRiddle", "IAmLordVoldemort") == True`
]

```{python, echo=FALSE, eval = FALSE}
def sortString(s):
    chars = sorted(s)
    return("".join(chars))

def areAnagrams(s1, s2):
    s1 = s1.lower()
    s2 = s2.lower()
    return(sortString(s1) == sortString(s2))
```
]

---
# [HW 8](https://p4a.seas.gwu.edu/2020-Fall/hw8-python.html)

I suggest starting with `reticulate::repl_python()` to work in Python from RStudio.

--
- Submit your "hw8.py" file to the autograder - it will (hopefully) work
