
class: inverse, middle

# Week 2: .fancy[Functions & Packages]

## 1. Functions
## 2. .orange[Manipulating data types]
## 3. Packages
## 4. Polya’s Problem Solving Technique

---
# Use these patterns:

--
.leftcol40[
### Convert type of `x`:
###`as.______(x)`
<br>

### Check type of `x`:
### `is.______(x)`
]
--
.rightcol60[
### Replace "`______`" with:

- ### `character`
- ### `logical`
- ### `numeric` / `double` / `integer`
]

---
## Convert type with `as.______(x)`

--
.leftcol[
### Convert **numeric** types:
```{r}
as.numeric("3.1415")
as.double("3.1415")
as.integer("3.1415")
```
]
--
.rightcol[
### Convert **non-numeric** types:
```{r}
as.character(3.1415)
as.logical(3.1415)
```
]

---
# A few notes on converting types

--
.leftcol[
### Converting any number to a logical returns `TRUE` except for `0`
```{r}
as.logical(7)
as.logical(0)
```
]
--
.rightcol[
### `TRUE = 1`, `FALSE = 0`:

```{r}
as.numeric(TRUE)
as.numeric(FALSE)
```
]

---
# A few notes on converting types

### Not everything can be converted.

--
.leftcol[
```{r}
as.numeric('7')   # Works
as.numeric('foo') # Doesn't work
```
]

---
# A few notes on converting types

### `as.integer()` is the same as `floor()`:
.leftcol[
```{r}
as.integer(3.14)
as.integer(3.99)
```
]

---
## Check type with `is.______(x)`

--
.leftcol[
### Checking **numeric** types:
```{r}
is.numeric(3.1415)
is.double(3.1415)
is.integer(3.1415)
```
]
--
.rightcol[
### Checking **non-numeric** types:
```{r}
is.character(3.1415)
is.logical(3.1415)
```
]

---
# Integers are weird

--
```{r}
is.integer(7)
```
...because R thinks `7` is really `7.0`
<br><br>

--
To check if a number is an integer _in value_:
```{r}
7 == as.integer(7)
```

---
class: inverse

# Think-Pair-Share

```{r, echo=FALSE}
countdown(minutes = 8,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```
Consider the following code (don't run it):
```{r, eval=FALSE}
number    <- as.logical(as.numeric('3'))
character <- is.character(typeof(7))
true      <- as.logical("FALSE")
false     <- as.logical(as.numeric(TRUE))

! (number == character) & (true | false) | (number & false)
```

Now follow these steps:

1. Type out what you expect R will return when all the lines are run at once.
2. Compare your expectations with each other.
3. Run the code and compare the results with your expectations.
