
class: inverse, middle

# Week 6: .fancy[Vectors]

## 1. .orange[Making vectors]
## 2. Vector operations
## 3. Comparing vectors
## 4. Slicing vectors
## 5. Lists

---
# Note: We've been using vectors already!

.code80[
```{r}
x <- 1
x
```
]
--
.code80[
```{r}
is.vector(x)
```
]
--
.code80[
```{r}
length(x)
```
]

---
# The universal vector generator: `c()`

--
.cols3[
## Numeric vectors
```{r}
x <- c(1, 2, 3)
x
```
]
--
.cols3[
## Character vectors
```{r}
y <- c('one', 'two', 'three')
y
```
]
--
.cols3[
## Logical vectors
```{r}
z <- c(TRUE, FALSE, TRUE)
z
```
]

---
# Elements in vectors must be the same type

### Type hierarchy:
- `character` > `numeric` > `logical`
- `double` > `integer`


--
.cols3[
Coverts to characters:
```{r}
c(1, "foo", TRUE)
```
]
--
.cols3[
Coverts to numbers:
```{r}
c(7, TRUE, FALSE)
```
]
--
.cols3[
Coverts to double:
```{r}
c(1L, 2, pi)
```
]

---
# Other ways to make a vector

--
.leftcol[
Sequences (we saw these last week):
```{r}
seq(1, 5)
```
```{r}
1:5
```
]
--
.rightcol[
Repeating a value:
```{r}
rep(5, 3)
```
```{r}
rep("snarf", 3)
```
]

---
# Repeating a vector

.leftcol[.code80[
Repeating a sequence
```{r}
x <- rep(seq(1, 3), 3)
x
length(x)
```
]]
--
.rightcol[.code80[
Note what the `each` argument does:
```{r}
x <- rep(seq(1, 3), each = 3)
x
length(x)
```
]]

---
# Quick code tracing

```{r, echo=FALSE}
countdown(minutes = 1,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

.code80[
What will each of these lines produce?
```{r, eval=FALSE}
rep(c(TRUE, FALSE, "TRUE"), 2)
```
```{r, eval=FALSE}
seq(FALSE, 3)
```
```{r, eval=FALSE}
rep(c(seq(3), seq(2)), each = 2)
```
]
