
class: inverse, middle

# Week 6: .fancy[Vectors]

## 1. Making vectors
## 2. .orange[Vector operations]
## 3. Slicing vectors
## 4. Comparing vectors
## 5. Lists

---
## Math on vectors is done **by element**

```{r}
x <- 1:10
```
--
```{r}
x + 2
```
--
```{r}
x - 2
```
--
```{r}
x * 2
```
--
```{r}
x / 2
```

---
## Math on vectors is done **by element**

--
```{r}
x1 <- c(1, 2, 3)
x2 <- c(4, 5, 6)
```
--
```{r}
x1 + x2 # Returns (1+4, 2+5, 3+6)
```
--
```{r}
x1 - x2 # Returns (1-4, 2-5, 3-6)
```
--
```{r}
x1 * x2 # Returns (1*4, 2*5, 3*6)
```
--
```{r}
x1 / x2 # Returns (1/4, 2/5, 3/6)
```

---
# If dimensions don't match, R "wraps" the vector

--
```{r}
x1 <- c(1, 2, 3, 4)
x2 <- c(4, 5)
```
```{r, eval=FALSE}
x1 + x2
```
--
```{r, echo=FALSE}
x1 + x2
```

--
```{r}
x1 <- c(1, 2, 3, 4)
x2 <- c(1)
```
```{r, eval=FALSE}
x1 + x2
```
--
```{r, echo=FALSE}
x1 + x2
```

---
# Most R functions work on vectors

--
.leftcol[.code80[
```{r}
x <- c(3.1415, 1.618, 2.718)
x
```
```{r}
round(x, 2)
```
```{r}
sqrt(x)
```
]]
--
.rightcol[.code80[
Works with your own functions too:
```{r}
isEven <- function(x) {
    return((x %% 2) == 0)
}
```
```{r}
x <- c(1, 4, 5, 10)
isEven(x)
```
]]

---
# Using vectors instead of a loop: **Summation**

Example: Sum the integers from 1 to 10
--
.leftcol[.code80[
Summing with a loop:
```{r}
x <- seq(1, 10)
total <- 0
for (item in x) {
    total <- total + item
}
total
```
]]
--
.rightcol[.code80[
Use a _summary function_ on the vector:
```{r}
sum(x)
```
]]

---
## "Summary" functions take a vector and return one value

--
```{r}
x <- 1:10
```
--
.leftcol[
```{r}
length(x)
sum(x)
prod(x)
```
]
--
.rightcol[
```{r}
min(x)
max(x)
mean(x)
median(x)
```
]

---
# Quick code tracing

```{r, echo=FALSE}
countdown(minutes = 3,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

.leftcol[.code80[
Consider this function:
```{r}
f <- function(x) {
    m <- x
    n <- sum(x + 4)
    m <- m + 5
    return(c(m, n))
}
```
]]
.rightcol[.code80[
What will this return?
```{r, eval=FALSE}
x <- c(1, 3)
f(x)
```
```{r, eval=FALSE}
y <- c(TRUE, FALSE, 1)
f(y)
```
]]
