
class: inverse, middle

# Week 6: .fancy[Vectors]

## 1. Making vectors
## 2. Vector operations
## 3. .orange[Slicing vectors]
## 4. Comparing vectors
## 5. Lists

---
# Use brackets `[]` to get elements from a vector

```{r}
x <- seq(1, 10)
```
--
.leftcol[
Indices start at 1:
```{r}
x[1] # Returns the first element
```
```{r}
x[3] # Returns the third element
```
```{r}
x[length(x)] # Returns the last element
```
]
--
.rightcol[
Slicing with a vector of indices:
```{r}
x[1:3]  # Returns the first three elements
```
```{r}
x[c(2, 7)] # Returns the 2nd and 7th elements
```
]

---
# Use negative integers to _remove_ elements

```{r}
x <- seq(1, 10)
```
--
```{r}
x[-1] # Drops the first element
```
--
```{r}
x[-1:-3] # Drops the first three elements
```
--
```{r}
x[-c(2, 7)] # Drops the 2nd and 7th elements
```
--
```{r}
x[-length(x)] # Drops the last element
```

---
# Slicing with logical indices

--
```{r}
x <- seq(1, 20, 3)
x
```
--
Create a logical vector based on some condition:
```{r}
x > 10
```
--
Slice `x` with logical vector - only `TRUE` indices will be returned:
```{r}
x[x > 10]
```

---
# You can also use `which()` to find indices

```{r}
x <- seq(1, 20, 3)
x
```
--
Use `which()` around a condition to get the indices where condition is `TRUE`:
```{r}
which(x > 10)
```
--
```{r}
x[which(x > 10)]
```

---
# You can name vector elements

```{r}
x <- seq(5)
x
```

--
.leftcol[
1) Add names with the `names()` function:
```{r}
names(x) <- c('a', 'b', 'c', 'd', 'e')
x
```
]
--
.rightcol[
2) Create a named vector:
```{r}
y <- c('a'=1, 'b'=2, 'c'=3, 'd'=4, 'e'=5)
y
```
]

---
# Using names to slice a vector

```{r}
x
```
--
```{r}
x['a']
```
--
```{r}
x[c('a', 'c')]
```

---
# Sorting vectors with `sort()`
--
```{r}
a = c(2, 4, 6, 3, 1, 5)
a
```
--
```{r}
sort(a)
```
--
```{r}
sort(a, decreasing = TRUE)
```

---
## `order()` returns the indices of the sorted vector

```{r}
a
order(a)
```

--
This does the same thing as `sort(a)`:
```{r}
a[order(a)]
```

---
# Quick code tracing

```{r, echo=FALSE}
countdown(minutes = 2,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

.leftcol[.code80[
Consider this function:
```{r, eval=FALSE}
f <- function(x) {
    for (i in seq(length(x))) {
        x[i] <- x[i] + sum(x) + max(x)
    }
    return(x)
}
```
]]
.rightcol[.code80[
What will this code return?
```{r, eval=FALSE}
x <- c(1, 2, 3)
f(x)
```
]]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

## Think-Pair-Share

.font80[
1) `reverse(x)`: Write a function that returns the vector in reverse order. You cannot use the `rev()` function.

- `reverse(c(5, 1, 3))` should return `c(3, 1, 5)`
- `reverse(c('a', 'b', 'c'))` should return `c('c', 'b', 'a')`
- `reverse(c(FALSE, TRUE, TRUE))` should return `c(TRUE, TRUE, FALSE)`

2) `alternatingSum(a)`: Write a function that takes a vector of numbers `a` and returns the alternating sum, where the sign alternates from positive to negative or vice versa.

- `alternatingSum(c(5,3,8,4))` should return `(5 - 3 + 8 - 4)`
- `alternatingSum(c(1,2,3))` should return `(1 - 2 + 3)`
- `alternatingSum(c(0,0,0))` should return `0`
- `alternatingSum(c(-7,5,3))` should return `(-7 - 5 + 3)`
]
