
class: inverse, middle

# Week 6: .fancy[Vectors]

## 1. Making vectors
## 2. Vector operations
## 3. Slicing vectors
## 4. .orange[Comparing vectors]
## 5. Lists

---
# Comparing vectors

Check if 2 vectors are the same:

.code80[
```{r}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
```
```{r, eval=FALSE}
x == y
```
]
--
.code80[
```{r, echo=FALSE}
x == y
```
]

---
# Comparing vectors with `all()` and `any()`

--
.leftcol[.code80[
`all()`: Check if _all_ elements are the same
```{r}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
all(x == y)
```
```{r}
x <- c(1, 2, 3)
y <- c(-1, 2, 3)
all(x == y)
```
]]
--
.rightcol[.code80[
`any()`: Check if _any_ elements are the same
```{r}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
any(x == y)
```
```{r}
x <- c(1, 2, 3)
y <- c(-1, 2, 3)
any(x == y)
```
]]

---
# `all()` vs. `identical()`

--
.code80[
```{r}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
names(x) <- c('a', 'b', 'c')
names(y) <- c('one', 'two', 'three')
```
]
--
.leftcol[.code80[
`all()` only compares the element _values_:
```{r}
all(x == y)
```
]]
--
.rightcol[.code80[
`identical()` compares _values_ and _names_:
```{r}
identical(x, y)
```
```{r}
names(y) <- names(x)
identical(x, y)
```
]]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

## Think-Pair-Share

Re-write `isPrime(n)` from last week, but **without loops!**. Remember, `isPrime(n)` takes a non-negative integer, `n`, and returns `TRUE` if it is a prime number and `FALSE` otherwise. Here's some test cases:

- `isPrime(1) == FALSE`
- `isPrime(2) == TRUE`
- `isPrime(7) == TRUE`
- `isPrime(13) == TRUE`
- `isPrime(14) == FALSE`

(If you're stuck, go to the next slide for a hint)

---
class: inverse

## Hint

Loop solution:
.code80[
```{r, eval=FALSE}
isPrime <- function(n) {
    if (n <= 1) { return(FALSE) }
    if (n == 2) { return(TRUE) }
    for (i in seq(2, (n-1))) {
        if ((n %% i) == 0) {
            return(FALSE)
        }
    }
    return(TRUE)
}
```
]
