
class: inverse, middle

# Week 6: .fancy[Vectors]

## 1. Making vectors
## 2. Vector operations
## 3. Comparing vectors
## 4. Slicing vectors
## 5. .orange[Lists]

---
# Lists are like vectors that can store anything

--
.leftcol[.code80[
Vectors force things to one type:
```{r}
c(1, "foo", TRUE)
```
]]
--
.rightcol[.code80[
Lists store any type:
```{r}
list(1, "foo", TRUE)
```
]]

---
# Elements in lists can be any object

--
.leftcol[.code80[
List of vectors:
```{r}
list(c(1, 2, 3), c("foo", "bar"), TRUE)
```
]]
--
.rightcol[.code80[
List of vector and function:
```{r}
vector <- c(1, 2, 3)
square <- function(x) {
    return(x^2)
}

list(vector, square)
```
]]

---
# Slice list with indices or names

.leftcol[.code80[
Slice with index using `[[]]`
```{r}
x <- list(
  c(1, 2, 3),
  c("foo", "bar"),
  TRUE)
```
```{r}
x[[1]]
x[[2]]
```
]]
--
.rightcol[.code80[
Slice with name using ``[[]]` or `$`
```{r}
x <- list(
  numbers = c(1, 2, 3),
  chars   = c("foo", "bar"),
  logical = TRUE)
```
```{r}
x[['numbers']]
x$numbers
```
]]

---
# [HW 6](https://p4a.seas.gwu.edu/2020-Fall/hw6-vectors.html)

- This is the last HW due before before the midterm (check the [schedule](https://p4a.seas.gwu.edu/2020-Fall/schedule.html)).

--
- Next week is Quiz 4 - also the last quiz before the midterm.

--
- Midterm is during class period on Oct. 20.


