
class: inverse, middle

# Week 3: .fancy[Creating Functions]

## 1. .orange[Function syntax]
## 2. Local vs global variables
## 3. Top-down design
## 4. Coding style

---
# Basic function syntax

.code90[
```{r eval=FALSE}
functionName <- function(arguments) {
    # Do stuff here
    return(something)
}
```
]

---
# Basic function syntax
In English:

> "`functionName` is a `function` of `arguments` that does..."

.code90[
```{r eval=FALSE}
functionName <- function(arguments) {
    # Do stuff here
    return(something)
}
```
]

---
# Basic function syntax
Example:

> "`squareRoot` is a `function` of `n` that...returns the square root of `n`"

.code90[
```{r}
squareRoot <- function(n) {
    return(n^0.5)
}
```
]
--
.code90[
```{r}
squareRoot(64)
```
]

---
# `return()` and `cat()` statements

--
.leftcol[.code90[
```{r}
isPositive <- function(n) {
    return(n > 0)
}
```
]]
--
.rightcol[.code90[
```{r}
isPositive <- function(n) {
    cat(n > 0)
}
```
]]

---
# `return()` and `cat()` statements

.leftcol[.code90[
```{r}
isPositive <- function(n) {
    return(n > 0)
}
```
`return()` _returns_ back a value
```{r, eval=FALSE}
test <- isPositive(7)
test
```
```
TRUE
```
]]
.rightcol[.code90[
```{r}
isPositive <- function(n) {
    cat(n > 0)
}
```
`cat()` _prints_ a value as a string
```{r, eval=FALSE}
test <- isPositive(7)
```
```
TRUE
```
```{r, eval=FALSE}
test
```
```
Error: object 'test' not found
```
]]

---
## `cat()` is short for "concatenating"

--
```{r}
print_x <- function(x) {
    cat("The value of x is", x)
}
```
--
```{r}
print_x(7)
```
--
```{r}
print_x_squared <- function(x) {
    cat("The value of x is", x, "and the value of x^2 is", x^2)
}
```
--
```{r}
print_x_squared(7)
```

---
## `cat()` adds a space between values by default

--
```{r}
print_x <- function(x) {
    cat("The value of x is", x)
}
```
--
```{r}
print_x(7)
```
--
Modify separator with the `sep` argument:
```{r}
print_x <- function(x) {
    cat("The value of x is", x, sep = ": ")
}
```
--
```{r}
print_x(7)
```

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 2,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

# Code tracing practice

.leftcol[.code80[
Consider these functions:
```{r}
f1 <- function(x) {
    return(x^3)
}
f2 <- function(x) {
    cat(x^3)
}
f3 <- function(x) {
    cat(x^3)
    return(x^4)
}
f4 <- function(x) {
    return(x^3)
    cat(x^4)
}
```
]]
.rightcol[.code80[
What will these lines of code produce?

Write your answer down first, _then_ run the code to check.
```{r, eval=FALSE}
f1(2)
f2(2)
f3(2)
f4(2)
```
]]
