
class: inverse, middle

# Week 7: .fancy[Strings]

## 1. Making strings
## 2. .orange[Case conversion & substrings]
## 3. Padding, splitting, & merging
## 4. Detecting & replacing

---
class: center

## Case conversion & substrings
|Function         |  Description                            |
|:----------------|:----------------------------------------|
|`str_to_lower()` | converts string to lower case           |
|`str_to_upper()` | converts string to upper case           |
|`str_to_title()` | converts string to title case           |
|`str_length()`   | number of characters                    |
|`str_sub()`      | extracts substrings                     |
|`str_locate()`   | returns indices of substrings           |
|`str_dup()`      | duplicates characters                   |

---
# Case conversion

```{r}
x <- "Want to hear a joke about paper? Never mind, it's tearable."
```
--
```{r}
str_to_lower(x)
```
--
```{r}
str_to_upper(x)
```
--
```{r}
str_to_title(x)
```

---
# Comparing strings

.leftcol[
Case matters:
```{r}
a <- "Apples"
b <- "apples"
a == b
```
]
--
.rightcol[
Convert case _before_ comparing if you just want to compare the string text:
```{r}
str_to_lower(a) == str_to_lower(b)
str_to_upper(a) == str_to_upper(b)
```
]

---
# Get the number of characters in a string

--
.leftcol[
The `length()` function returns the _vector_ length:
```{r}
length("hello world")
```
]
--
.rightcol[
To get the # of characters, use `str_length()`:
```{r}
str_length("hello world")
str_length(" ") # Spaces count
str_length("")  # Empty string
```
]

---
## Access characters by their index with `str_sub()`

--
.leftcol[
Indices start at 1:
```{r}
str_sub("Apple", 1, 3)
```
Negative numbers count backwards from end:
```{r}
str_sub("Apple", -3, -1)
```
]
--
.rightcol[
Modify a string with `str_sub()`:
```{r}
x <- 'abcdef'
str_sub(x, 1, 3) <- 'ABC'
x
```
]

---

## Get the indices of substrings

Extract the substring `"Good"` from the following string:
```{r}
x <- 'thisIsGoodPractice'
```
--
.leftcol[
**1)**: Use `str_locate()` to get the **start** and **end** indices:
```{r}
indices <- str_locate(x, 'Good')
indices
```
]
--
.rightcol[
**2)**: Use `str_sub()` to get the substring:
```{r}
str_sub(x, indices[1], indices[2])
```
]

---
# Repeat a string with `str_dup()`

```{r}
str_dup("holla", 3)
```
--
Note the difference with `rep()`:
```{r}
rep("holla", 3)
```

---
# `stringr` functions work on vectors

--
```{r}
x <- c("apples", "oranges")
x
```
--
.leftcol[
Get the first 3 letters in each string:
```{r}
str_sub(x, 1, 3)
```
]
--
.rightcol[
Duplicate each string twice
```{r}
str_dup(x, 2)
```
]

---
# Quick practice:

```{r, echo=FALSE}
countdown(minutes = 5,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

Create this string object: `x <- 'thisIsGoodPractice'`, then use **stringr** functions to transform `x` into the following strings:

.leftcol[
- `'thisIsGood'`
- `'practice'`
- `'GOOD'`
- `'thisthisthis'`
- `'GOODGOODGOOD'`
]
.rightcol[
**Hint**: You'll need these:

- `str_to_lower()`
- `str_to_upper()`
- `str_locate()`
- `str_sub()`
- `str_dup()`
]
