
class: inverse, middle

# Week 7: .fancy[Strings]

## 1. Making strings
## 2. Case conversion & substrings
## 3. Padding, splitting, & merging
## 4. .orange[Detecting & replacing]


---
class: center

## Detecting & replacing
|Function         |  Description                            |
|:----------------|:----------------------------------------|
|`str_sort()`     | sort a string alphabetically            |
|`str_order()`    | get the order of a sorted string        |
|`str_detect()`   | match a string in another string        |
|`str_replace()`  | replace a string in another string      |

---
## Sort string vectors alphabetically with `str_sort()`

```{r}
x <- c('Y', 'M', 'C', 'A')
x
```
--
```{r}
str_sort(x)
```
--
```{r}
str_sort(x, decreasing = TRUE)
```

---
### Detect pattern in string: `str_detect(string, pattern)`

--
```{r}
tenFruit <- fruit[1:10]
tenFruit
```
--
```{r}
str_detect(tenFruit, "berry")
```
--
How many in vector have the string `"berry"`?
```{r}
sum(str_detect(tenFruit, "berry"))
```

---
## Count number of times pattern appears in string

`str_count(string, pattern)`
--
Example:
```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```
--
Note the difference with `str_detect()`:
```{r}
str_detect(x, "a")
```

---
## Detect if string _starts_ with pattern

Which fruits _start_ with "a"?
```{r}
fiveFruit <- fruit[1:5]
fiveFruit
```
--
.leftcol[
**Wrong**:
```{r}
str_detect(fiveFruit, "a")
```
]
--
.rightcol[
**Right**:
```{r}
str_detect(fiveFruit, "^a")
```
]

---
# Detect if string _ends_ with pattern

Which fruits _end_ with an "e"?
```{r}
fiveFruit
```
--
.leftcol[
**Wrong**:
```{r}
str_detect(fiveFruit, "e")
```
]
--
.rightcol[
**Right**:
```{r}
str_detect(fiveFruit, "e$")
```
]

---
## Remember:

### If you _start_ with power (`^`), you'll _end_ up with money (`$`).
--
```{r}
fiveFruit
```
--
```{r}
str_detect(fiveFruit, "^a") # Start with power (^)
```
--
```{r}
str_detect(fiveFruit, "e$") # End with money ($)
```

---
# Quick practice:

```{r, echo=FALSE}
countdown(minutes = 5,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

```{r}
fruit[1:5]
```

Use `stringr` functions to answer the following questions about the  `fruit` vector:

1. How many fruit have the string `"rr"` in it?
2. Which fruit end with string `"fruit"`?
3. Which fruit contain more than one `"o"` character?

**Hint**: You'll need to use `str_detect()` and `str_count()`

---
# Replace matched strings with new string

`str_replace(string, pattern, replacement)`
--
```{r}
x <- c("apple", "pear", "banana")
```
--
```{r}
str_replace(x, "a", "-") # Only replaces the first match
```
--
```{r}
str_replace_all(x, "a", "-") # Replaces all matches
```

---
### Quick practice redux

```{r}
x <- 'this_is_good_practice'
```
Convert `x` into: `"this is good practice"`

--

We did this earlier:
```{r}
paste(str_split(x, "_")[[1]], collapse = " ")
```

--
But now we can do this!
```{r}
str_replace_all(x, "_", " ")
```

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

## Think-Pair-Share

.font90[
1) `sortString(s)`: Write the function `sortString(s)` that takes a string `s` and returns back an alphabetically sorted string.

- `sortString("cba") == "abc"`
- `sortString("abedhg") == "abdegh"`
- `sortString("AbacBc") == "aAbBcc"`

2) `areAnagrams(s1, s2)`: Write the function `areAnagrams(s1, s2)` that takes two strings, `s1` and `s2`, and returns `TRUE` if the strings are [anagrams](https://en.wikipedia.org/wiki/Anagram), and `FALSE` otherwise. **Treat lower and upper case as the same letters**.

- `areAnagrams("", "") == TRUE`
- `areAnagrams("aabbccdd", "bbccddee") == FALSE`
- `areAnagrams("TomMarvoloRiddle", "IAmLordVoldemort") == TRUE`
]

---
### Homeworks

- This is the last week to submit homeworks 1 - 6


--
### Midterm

- Midterm is during class period next week, Oct. 20.
- Questions will be emailed as a PDF, type responses in .R file.


--
### Midterm Review

- Planning on holding a one-hour review this week (tentatively Thursday).
