
# R as a calculator

.leftcol[
### Basic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
]
--
.rightcol[
### Other important operators:

- Power: `^`
- Integer Division: `%/%`
- Modulus: `%%`
]

---
# Integer division: `%/%`

Integer division drops the remainder
<br><br>

--
&zwj;Example:
```{r}
4 / 3 # Regular division
4 %/% 3 # Integer division
```

---
# Integer division: `%/%`

Integer division drops the remainder<br><br>

--
&zwj;What will this return?
```{r, eval=FALSE}
4 %/% 4
```
--
```{r, echo=FALSE}
4 %/% 4
```

--
What will this return?
```{r, eval=FALSE}
4 %/% 5
```
--
```{r, echo=FALSE}
4 %/% 5
```

---
# Modulus operator: `%%`

Modulus returns the remainder _after_ doing integer division<br><br>

--
&zwj;Example:
```{r}
5 %% 3
```
<br>

--
```{r}
3.1415 %% 3
```

---
# Modulus operator: `%%`

Modulus returns the remainder _after_ doing integer division<br><br>

--
&zwj;What will this return?
```{r, eval=FALSE}
4 %% 4
```
--
```{r, echo=FALSE}
4 %% 4
```
--
What will this return?
```{r, eval=FALSE}
4 %% 5
```
--
```{r, echo=FALSE}
4 %% 5
```

---
## Odds and evens with `n %% 2`

--
.leftcol[
If `n %% 2` is `0`, `n` is **EVEN**
```{r}
10 %% 2
12 %% 2
```
Also works with negative numbers!
```{r}
-42 %% 2
```
]
--
.rightcol[
If `n %% 2` is `1`, `n` is **ODD**
```{r}
1 %% 2
13 %% 2
```
Also works with negative numbers!
```{r}
-47 %% 2
```
]

---
## Number "chopping" with 10s

--
.pull-left[
The mod operator (`%%`) "chops" a number and returns everything to the _right_

```{r}
123456 %% 1
123456 %% 10
123456 %% 100
```
]
--
.pull-right[
Integer division (`%/%`) "chops" a number and returns everything to the _left_

```{r}
123456 %/% 1
123456 %/% 10
123456 %/% 100
```
]

---
## Number "chopping" with 10s

- `%%` returns everything to the _right_ (`"chop" ->`)
- `%/%` returns everything to the _left_ (`<- "chop"`)
- The "chop" point is always just to the _right_ of the chopping digit:

<div style="width:1000px">
<table class="table table-condensed">
<thead>
<tr class="header">
<th>Example</th>
<th>“Chop” point</th>
<th>“Chop” point description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td><code>1234 %% 1</code></td>
<td><code>1234 |</code></td>
<td>Right of the <code>1</code>’s digit</td>
</tr>
<tr class="even">
<td><code>1234 %% 10</code></td>
<td><code>123 | 4</code></td>
<td>Right of the <code>10</code>’s digit</td>
</tr>
<tr class="odd">
<td><code>1234 %% 100</code></td>
<td><code>12 | 34</code></td>
<td>Right of the <code>100</code>’s digit</td>
</tr>
<tr class="even">
<td><code>1234 %% 1000</code></td>
<td><code>1 | 234</code></td>
<td>Right of the <code>1,000</code>’s digit</td>
</tr>
<tr class="odd">
<td><code>1234 %% 10000</code></td>
<td><code>| 1234</code></td>
<td>Right of the <code>10,000</code>’s digit</td>
</tr>
</tbody>
</table>
</div>

---
# Comparing things: **Relational operators**

--
.leftcol[
### Compare if condition is `TRUE` or `FALSE` using:

- Less than: `<`
- Less than or equal to : `<=`
- Greater than or equal to: `>=`
- Greater than: `>`
- Equal: `==`
- Not equal: `!=`
]
--
.rightcol[.code60[
```{r}
2 < 2
2 <= 2
(2 + 2) == 4
(2 + 2) != 4
"penguin" == "penguin"
```
]]

---
# Comparing things: **Logical operators**

--
.leftcol[
### Make multiple comparisons with:

- And: `&`
- Or: `|`
- Not: `!`
]
--
.rightcol[.code60[
With "and" (`&`), every part must be `TRUE`, otherwise the whole statement is `FALSE`:
```{r}
(2 == 2) & (3 == 3)
(2 == 2) & (2 == 3)
```
]]

---
# Comparing things: **Logical operators**

.leftcol[
### Make multiple comparisons with:

- And: `&`
- Or: `|`
- Not: `!`
]
.rightcol[.code60[
With "or" (`|`), if _any_ part is `TRUE`, the whole statement is `TRUE`:
```{r}
(2 == 2) | (3 == 3)
(2 == 2) | (2 == 3)
```
]]

---
# Comparing things: **Logical operators**

.leftcol[
### Make multiple comparisons with:

- And: `&`
- Or: `|`
- Not: `!`
]
.rightcol[.code60[
The "not" (`!`) symbol produces the _opposite_ statement:
```{r}
! (2 == 2)
! ((2 == 2) | (2 == 3))
```
]]

---
# Comparing things: **Logical operators**

.cols3[
### And: `&`
"Are any of the statements `FALSE`"?
```{r}
(2 == 2) & (2 == 3) & (4 == 4)
```
]
--
.cols3[
### Or: `|`
"Are any of the statements `TRUE`"?
```{r}
(2 == 2) | (2 == 3) | (4 == 7)
```
]
--
.cols3[
### `& > |`
The `&` operator takes precedence over `|`
```{r}
(2 == 2) | (2 == 3) & (4 == 7)
```
]

---
# Comparing things: **Logical operators**

.leftcol[
### **Pro tip**: Use parentheses

```{r}
!3 == 5   # Confusing
!(3 == 5) # Less confusing
```
]

---
## Other important points

.leftcol[
### R follows BEDMAS:

1. **B**rackets
2. **E**xponents
3. **D**ivision
4. **M**ultiplication
5. **A**ddition
6. **S**ubtraction
]
--
.rightcol[
### **Pro tip**: Use parentheses
```{r}
1 + 2 * 4   # Confusing
1 + (2 * 4) # Less confusing
```
]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 10,
          warn_when = 30,
          update_every = 15,
          top = 0,
          font_size = '2em')
```

# Your turn

.font80[
Consider the following objects:

```{r}
w <- TRUE
x <- FALSE
y <- TRUE
```

Write code to answer the following questions:
1. Fill in _relational_ operators to make the following statement return `TRUE`:

    `! (w __ x) & ! (y __ x)`

2. Fill in _logical_ operators to make this statement return `FALSE`:

    `! (w __ x) | (y __ x)`
]

```{r, eval=FALSE, echo=FALSE}
# 1.  Use relational operators to return TRUE
! (w == x) & ! (y == x)

# 2. Use logical operators to return FALSE
! (w | x) | (y & x)
```

---
# Data Types

|Type      | Description                 | Example
|:---------|:----------------------------|:---------
|`double`   | Numbers w/decimals (aka "float") | `3.14`
|`integer`   | Numbers w/out decimals    | `42`
|`character` | Text (aka "string") | `"this is some text"`
|`logical`   | Used for comparing objects | `TRUE`, `FALSE`

--
Use `typeof()` to assess the type of any variable:
```{r}
typeof("hello")
```

---
# Numeric types (there are 2)

--
.leftcol[
## Integers

No decimals (e.g. `7`)
]
--
.rightcol[
## Doubles (aka float")

Decimals (e.g. `7.0`)
]

---
## In R, numbers are "doubles" by default

--
&zwj;Example:
```{r}
typeof(3)
```
Even though it _looks_ like an integer, R assumes that `3` is really `3.0`
<br><br>

--
Make it an integer by adding `L`:
```{r}
typeof(3L)
```

---
# Character types

--
### Use single or double quotes around anything:

```{r}
typeof('hello')
typeof("3")
```

--
Use single / double quotes if the string _contains_ a quote symbol:
```{r}
typeof("don't")
```

---
# Logical types

--
.leftcol[
Logical data only have two values:<br>`TRUE` or `FALSE`
```{r}
typeof(TRUE)
typeof(FALSE)
```
]
--
.rightcol[
Note that these have to be in all caps,<br>and **not** in quotes:
```{r, error=TRUE}
typeof('TRUE')
typeof(True)
```
]
---
# Logical types

Use to answer questions about logical statements.

Example: Is `1` greater than `2`?
```{r}
1 > 2
```

--
&zwj;Example: Is `2` greater than `1`?
```{r}
1 < 2
```

---
# Special values

<!-- Adjust grid size...use 100 if no header -->
<style>
.wrapper {
  width: 70vw;
  height: 60vh;
}
</style>

.wrapper[
.top-left[
**Infinity**: `Inf`<br>_really big numbers_
```{r}
1/0
```
]
.top-right[
**Not a Number**: `NaN`<br>_"not a number"_
```{r}
0/0
```
]
.bottom-left[
**Not available**: `NA`<br>_value is "missing"_
]
.bottom-right[
**No value**: `NULL`<br>_no value whatsoever_
]]

---
class: inverse

# Your turn

```{r, echo=FALSE}
countdown(minutes = 5,
          warn_when = 30,
          update_every = 15,
          top = 0,
          font_size = '2em')
```

## Will these return `TRUE` or `FALSE`?

## (try to answer first, then run the code to check)

- `! typeof('3') == typeof(3)`
- `(typeof(7) != typeof("FALSE")) | FALSE`
- `! (typeof(TRUE) == typeof(FALSE)) & FALSE`

---
# Final points
<br>

--
## 1) [HW 1](https://p4a.seas.gwu.edu/2020-Fall/hw1-getting-started.html) Preview

## `r fa('exclamation-circle')` Read carefully!
<br>

--
## 2) Please take this [survey](https://p4aintrosurvey.formr.org/)
