
class: inverse, middle

# Week 11: .fancy[Data Frames]

## 1. Basics
## 2. .orange[Slicing]
## 3. External data

---
## &zwj; Access data frame columns using the `$` symbol

```{r, echo=FALSE}
beatles <- beatles_reset
```

```{r}
beatles$firstName
```
--
```{r}
beatles$lastName
```

---
# Creating new variables with the `$` symbol

--
Add the hometown of the bandmembers:
```{r}
beatles$hometown <- 'Liverpool'
beatles
```

---
# Creating new variables with the `$` symbol

--
Add a new `alive` variable:
```{r}
beatles$alive <- c(FALSE, TRUE, TRUE, FALSE)
beatles
```

---
## You can compute new variables from current ones

--
Compute and add the age of the bandmembers:
```{r}
beatles$age <- 2020 - beatles$yearOfBirth
beatles
```

---
## Access elements by index: `DF[row, column]`

```{r, echo=FALSE}
beatles <- tibble(
    firstName   = c("John", "Paul", "Ringo", "George"),
    lastName    = c("Lennon", "McCartney", "Starr", "Harrison"),
    instrument  = c("guitar", "bass", "drums", "guitar"),
    yearOfBirth = c(1940, 1942, 1940, 1943),
    deceased    = c(TRUE, FALSE, FALSE, TRUE)
)
```

General form for indexing elements:
```{r, eval=FALSE}
DF[row, column]
```
--
.leftcol[
Select the element in row 1, column 2:
```{r}
beatles[1, 2]
```
]
--
.rightcol[
Select the elements in rows 1 & 2 and columns 2 & 3:
```{r}
beatles[c(1, 2), c(2, 3)]
```
]

---
## Leave row or column "blank" to select all

--
```{r}
beatles[c(1, 2),] # Selects all COLUMNS for rows 1 & 2
```
--
```{r}
beatles[,c(1, 2)] # Selects all ROWS for columns 1 & 2
```

---
## Negative indices exclude row / column

--
```{r}
beatles[-1, ] # Select all ROWS except the first
```
--
```{r}
beatles[,-1] # Select all COLUMNS except the first
```

---
# You can select columns by their names

Note: you don't need the comma to select an entire column

--
.leftcol[
One column
```{r}
beatles['firstName']
```
]
--
<br>Multiple columns
.rightcol[
```{r}
beatles[c('firstName', 'lastName')]
```
]

---
## Use logical indices to _filter_ rows

--
**Which Beatles members are still alive?**<br>Create a logical vector using the `deceased` column:
```{r}
beatles$deceased == FALSE
```
--
Insert this logical vector in the ROW position of `beatles[,]`:
```{r}
beatles[beatles$deceased == FALSE,]
```

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 10,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

## Think-Pair-Share

Answer these questions using the `beatles` data frame:

1. Create a new column, `playsGuitar`, which is `TRUE` if the band member plays the guitar and `FALSE` otherwise.
2. Filter the data frame to select only the rows for the band members who have four-letter first names.
3. Create a new column, `fullName`, which contains the band member's first and last name separated by a space (e.g. `"John Lennon"`)

```{r, echo=FALSE, eval=FALSE}
beatles <- tibble(
    firstName   = c("John", "Paul", "Ringo", "George"),
    lastName    = c("Lennon", "McCartney", "Starr", "Harrison"),
    instrument  = c("guitar", "bass", "drums", "guitar"),
    yearOfBirth = c(1940, 1942, 1940, 1943),
    deceased    = c(TRUE, FALSE, FALSE, TRUE)
)
beatles$playsGuitar <- beatles$instrument == 'guitar'
library(stringr)
beatles[str_length(beatles$firstName) == 4,]
beatles$fullName <- paste(beatles$firstName, beatles$lastName, sep = ' ')
```
