
class: inverse, middle

# Week 13: .fancy[Data Visualization]

## 1. .orange[Plotting with Base R]
## 2. Plotting with **ggplot2**
## 3. Tweaking your ggplot

---
# Today's data:
## Bear attacks in North America

Explore the `bears` data frame:
```{r, eval=FALSE}
glimpse(bears)
head(bears)
```

---
class: center

## Two basic plots in R

.leftcol[
### Scatterplots
```{r, base_scatter, echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
plot(x    = bears$year,
     y    = bears$age,
     col  = 'darkblue', # "col" changes the point color
     pch  = 19, # "pch" changes the point shape
     main = "Age of bear killing victims over time",
     xlab = "Year",
     ylab = "Age")
box()
```
]
.rightcol[
### Histograms
```{r, base_hist, echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE, class = "thin-outline"}
hist(x      = bears$month,
     breaks = 12,
     col    = 'darkred',
     main   = "Distribution of bear killings by month",
     xlab   = "Month",
     ylab   = "Count")
```
]

---
# Scatterplots with `plot()`

### Plot relationship between two variables

--
.leftcol[
General syntax:
```{r, eval=FALSE}
plot(x = x_vector, y = y_vector)
```
]

---
# Scatterplots with `plot()`

### Plot relationship between two variables

.leftcol[
General syntax:
```{r, eval=FALSE}
plot(x = x_vector, y = y_vector)
```

Example:
```{r scatterplot, fig.show='hide'}
var1 <- seq(1, 5)
var2 <- 2*var1
plot(x = var1, y = var2)
```
]
.rightcol[
```{r ref.label='scatterplot', echo=FALSE, fig.height=5, fig.width=6}
```
]

---
# Scatterplots with `plot()`

### `x` and `y` must have the same length!

--
```{r}
var2 <- var2[-1]
```
--
```{r}
length(var1) == length(var2)
```
--
```{r, message=FALSE, warning=FALSE, error=TRUE}
plot(x = var1, y = var2)
```

---
# Scatterplots with `plot()`

### Plotting variables from a data frame:

--
.leftcol[
Plot `year` vs. `age`:
```{r bear_scatterplot, fig.show='hide'}
plot(x = bears$year, y = bears$age)
```
]
--
.rightcol[
```{r ref.label='bear_scatterplot', echo=FALSE, fig.height=5, fig.width=7}
```
]

---
# Making `plot()` pretty

.leftcol[.code80[
```{r bear_prettyPlot, fig.show='hide'}
plot(x    = bears$year,
     y    = bears$age,
     col  = 'darkblue', # Point color
     pch  = 19, # Point shape
     main = "Age of victims over time",
     xlab = "Year",
     ylab = "Age of victim")
```
]]
.rightcol[
```{r ref.label='bear_prettyPlot', echo=FALSE, fig.height=5, fig.width=7}
```
]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 10,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

## Think pair share: `plot()`

Does the annual number of bird impacts appear to be changing over time?

Make a plot using the `birds` data frame to justify your answer

Hint: You may need to create a summary data frame to answer this question!

**Bonus points**: Make your plot pretty

---
# Histograms with `hist()`

### Plot the _distribution_ of a single variable

.leftcol[
General syntax:
```{r, eval=FALSE}
hist(x = x_vector)
```
]

---
# Histograms with `hist()`

### Plot the _distribution_ of a single variable

.leftcol[
General syntax:
```{r, eval=FALSE}
hist(x = x_vector)
```

Example:
```{r bearHist, fig.show='hide'}
hist(bears$month)
```
]
.rightcol[
```{r ref.label='bearHist', echo=FALSE, fig.height=5, fig.width=6}
```
]

---
# Making `hist()` pretty

.leftcol[.code80[
```{r prettyHist, fig.show='hide'}
hist(x      = bears$month,
     breaks = 12,
     col    = 'darkred',
     main   = "Bear killings by month",
     xlab   = "Month",
     ylab   = "Count")
```
]]
.rightcol[
```{r ref.label='prettyHist', echo=FALSE, fig.height=5, fig.width=6}
```
]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 10,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

## Think pair share: `hist()`

Make plots using the `birds` data frame to answer these questions

- Which months have the highest and lowest number of bird impacts in the dataset?
- Which aircrafts experience more impacts: 2-engine, 3-engine, or 4-engine?
- At what height do most impacts occur?

**Bonus points**: Make your plots pretty
