## Make bar charts with `geom_col()`

With bar charts, you'll often need to create summary variables to plot

--
.leftcol[
Step 1: Summarize the data
```{r}
bear_months <- bears %>%
  count(month) #<<
```

Step 2: Make the plot

```{r bears_geom_col, fig.show='hide'}
ggplot(bear_months) +
  geom_col(aes(x = month, y = n)) #<<
```
]
.rightcol[
Example: count of attacks by month
```{r ref.label='bears_geom_col', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---
## Make bar charts with `geom_col()`

Alternative approach: piping directly into ggplot
.leftcol[
```{r bears_geom_col_pipe, fig.show='hide'}
bears %>%
  count(month) %>%  # Pipe into ggplot #<<
  ggplot() +
  geom_col(aes(x = month, y = n))
```
]
.rightcol[
```{r ref.label='bears_geom_col_pipe', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---
## Be careful with `geom_col()` vs. `geom_bar()`

.leftcol[
### .center[`geom_col()`]

Map both `x` and `y`

```{r, fig.height=3, fig.width=4}
bears %>%
  count(month) %>% #<<
  ggplot() +
  geom_col(aes(x = month, y = n)) #<<
```
]
.rightcol[
### .center[`geom_bar()`]

Only map `x` (`y` is computed)

```{r, fig.height=3, fig.width=4}
bears %>%
  ggplot() +
  geom_bar(aes(x = month)) #<<
```
]

---
## Make bar charts with `geom_col()`

.leftcol[
Another example:<br>Mean age of victim in each year

```{r bears_mean_age, fig.show='hide'}
bears %>%
  filter(!is.na(age)) %>%
  group_by(year) %>%
  summarise(meanAge = mean(age)) %>% #<<
  ggplot() +
  geom_col(aes(x = year, y = meanAge)) #<<
```
]
.rightcol[.blackborder[
```{r ref.label='bears_mean_age', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---
### Change bar width:  `width`
### Change bar color: `fill`
### Change bar outline: `color`

.leftcol[
```{r bears_geom_col_colors, fig.show='hide'}
bears %>%
  count(month) %>%
  ggplot() +
  geom_col(aes(x = month, y = n),
           width = 0.7, #<<
           fill = "blue", #<<
           color = "red") #<<
```
]
.rightcol[
```{r ref.label='bears_geom_col_colors', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---
### Map the `fill` to `bearType`

.leftcol[.code70[
```{r bears_geom_col_fill, fig.show='hide'}
bears %>%
  count(month, bearType) %>% #<<
  ggplot() +
  geom_col(aes(x = month, y = n,
               fill = bearType)) #<<
```
]

Note that I had to summarize the count by both `month` and `bearType`

.code70[
```{r, eval=FALSE}
bears %>%
  count(month, bearType)
```
]
.code60[
```{r, echo=FALSE}
bears %>%
  count(month, bearType) %>%
  arrange(month, bearType)
```
]]
.rightcol[
```{r ref.label='bears_geom_col_fill', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---
# "Factors" = Categorical variables

By default, R makes numeric variables _continuous_

.leftcol[
```{r, eval=FALSE}
bears %>%
  count(month) %>%
  ggplot() +
  geom_col(aes(x = month, y = n)) 
```
**The variable `month` is a _number_**
]
.rightcol[.blackborder[
```{r ref.label='bears_geom_col_pipe', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]]

---
# "Factors" = Categorical variables

You can make a continuous variable _categorical_ using `as.factor()`

.leftcol[.code80[
```{r bears_geom_col_factor, fig.show='hide'}
bears %>%
  count(month) %>%
  ggplot() +
  geom_col(aes(x = as.factor(month), #<<
               y = n)) 
```
]

**The variable `month` is a _factor_**
]
.rightcol[.blackborder[
```{r ref.label='bears_geom_col_factor', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

## Think pair share: `geom_col()`

Use the `bears` and `birds` data frame to create the following plots

.leftcol[
```{r, ggbar_p1, echo=FALSE, fig.height=5, fig.width=7}
bears %>% 
  count(year, gender) %>% 
  ggplot() +
  geom_col(aes(x = year, y = n, fill = gender)) +
  labs(x     = "Year",
       y     = 'Number of killings',
       fill  = "Victim gender",
       title = "Annual deadly bear attacks over time")
```
]
.rightcol[
```{r, ggbar_p2, echo=FALSE, fig.height=5, fig.width=7}
birds %>%
  filter(!is.na(cost_repairs_infl_adj)) %>%
  group_by(incident_year) %>%
  summarise(maxCost = max(cost_repairs_infl_adj)) %>% 
  ggplot() +
  geom_col(aes(x = incident_year, y = maxCost)) +
  labs(x     = "Year",
       y     = 'Maximum cost ($)',
       title = "Max annual cost from wildlife impact")
```
]
