class: inverse, middle

# Week 13: .fancy[Data Visualization]

## 1. Plotting with Base R
## 2. Plotting with **ggplot2**
## 3. .orange[Tweaking your ggplot]

---
# Working with themes

Themes change _global_ features of your plot, like the background color, grid lines, etc.

--
.leftcol[
```{r ggbears_basic, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point()
```
]
.rightcol[
```{r ref.label='ggbears_basic', echo=FALSE, fig.height=5, fig.width=7}
```
]

---
# Working with themes

Themes change _global_ features of your plot, like the background color, grid lines, etc.

.leftcol[
```{r ggbears_theme, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_bw() #<<
```
]
.rightcol[
```{r ref.label='ggbears_theme', echo=FALSE, fig.height=5, fig.width=7}
```
]

---
### Common themes

.leftcol[
`theme_bw()`
```{r, fig.height=4, fig.width=6}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_bw() #<<
```
]
.rightcol[
`theme_minimal()`
```{r, fig.height=4, fig.width=6}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_minimal() #<<
```
]

---
### Common themes

.leftcol[
`theme_classic()`
```{r, fig.height=4, fig.width=6}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_classic() #<<
```
]
.rightcol[
`theme_void()`
```{r, fig.height=4, fig.width=6}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_void() #<<
```
]

---
### Other themes: [**hrbrthemes**](https://github.com/hrbrmstr/hrbrthemes)

.leftcol[
```{r, fig.height=4, fig.width=6}
library(hrbrthemes) #<<

ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_ipsum() #<<
```
]
.rightcol[
```{r, fig.height=4, fig.width=6}
library(hrbrthemes) #<<

ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_ft_rc() #<<
```
]

---
### Other themes: **ggthemes**

.leftcol[
```{r, fig.height=4, fig.width=6}
library(ggthemes) #<<

ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_economist() #<<
```
]
.rightcol[
```{r, fig.height=4, fig.width=6}
library(ggthemes) #<<

ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_economist_white() #<<
```
]

---
# Save figures with `ggsave()`

--
First, assign the plot to an object name:
```{r, eval=FALSE}
scatterPlot <- ggplot(data = bears) +
    geom_point(aes(x = year, y = age))
```

--
Then use `ggsave()` to save the plot:
```{r, eval=FALSE}
ggsave(filename = here('plots', 'scatterPlot.png'),
       plot   = scatterPlot,
       width  = 6, # inches
       height = 4)
```

---
class: inverse

# Extra practice 1

Use the `mtcars` data frame to create the following plots

.cols3[
```{r, mtcars_1, echo=FALSE, fig.height=4, fig.width=4.5}
ggplot(mtcars, aes(x=wt, y=mpg)) +
    geom_point() +
    theme_bw() +
    labs(x = "Weight",
         y = 'Fuel Economy (mpg)')
```
]
.cols3[
```{r, mtcars_2, echo=FALSE, fig.height=4, fig.width=4.5}
ggplot(mtcars, aes(x=hp, y=mpg)) +
    geom_point(aes(color=as.factor(cyl))) +
    theme_bw() +
    labs(x = "Horsepower (hp)",
         y = 'Fuel Economy (mpg)',
         color = 'Num. Cylinders')
```
]
.cols3[
```{r, mtcars_3, echo=FALSE, fig.height=4, fig.width=4.5}
mtcars %>%
  group_by(cyl) %>%
  summarise(meanMpg = mean(mpg)) %>%
  ggplot(aes(x=cyl, y=meanMpg)) +
    geom_bar(stat = 'identity') +
    labs(x = "Num. Cylinders",
         y = 'Mean Fuel Economy (mpg)')
```
]

---
class: inverse

# Extra practice 2

Use the `mpg` data frame to create the following plot

```{r, mtcars_4, echo=FALSE, fig.height=6, fig.width=8}
mpg %>%
    mutate(manufacturer = str_to_title(manufacturer)) %>%
    group_by(manufacturer) %>%
    summarise(meanHwy = mean(hwy)) %>%
    mutate(manufacturer = fct_reorder(manufacturer, meanHwy)) %>%
    ggplot(aes(x=manufacturer, y=meanHwy)) +
    geom_bar(stat='identity') +
    coord_flip() +
    labs(
        x='Vehicle manufacturer',
        y='Highway fuel economy (mpg)')
```
