
class: inverse, middle

# Week 4: .fancy[Conditionals & Testing]

## 1. Conditionals
## 2. .orange[Testing]
## 3. Tips

---
# Why write test functions?
<br>

--
## 1. They help you understand the problem

--
## 2. They verify that a function is working as expected

---
# Test function "syntax"

--
.leftcol[.code80[
### Function:
```{r, eval=FALSE}
functionName <- function(arguments) {
    # Do stuff here
    return(something)
}
```
]]
--
.rightcol[.code80[
### Test function:
```{r, eval=FALSE}
test_functionName <- function() {
    cat("Testing functionName()...")
    # Put test cases here
    cat("Passed!\n")
}
```
]]

---
# Writing test cases with `stopifnot()`

`stopifnot()` stops the function if whatever is inside the `()` is not `TRUE`.

--
.leftcol[.code80[
### Function:
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
- `isEven(1)` should be `FALSE`
- `isEven(2)` should be `TRUE`
- `isEven(-7)` should be `FALSE`
]]
--
.rightcol[.code80[
### Test function:
```{r}
test_isEven <- function() {
    cat("Testing isEven()...")
    stopifnot(isEven(1) == FALSE)
    stopifnot(isEven(2) == TRUE)
    stopifnot(isEven(-7) == FALSE)
    cat("Passed!\n")
}
```
]]

---
# Writing test cases with `stopifnot()`

`stopifnot()` stops the function if whatever is inside the `()` is not `TRUE`.

.leftcol[.code80[
### Function:
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
- `isEven(1)` should be `FALSE`
- `isEven(2)` should be `TRUE`
- `isEven(-7)` should be `FALSE`
]]
.rightcol[.code80[
### Test function:
```{r}
test_isEven <- function() {
    cat("Testing isEven()...")
    stopifnot(isEven(1) == FALSE)
    stopifnot(isEven(2) == TRUE)
    stopifnot(isEven(-7) == FALSE)
    cat("Passed!\n")
}
```
```{r}
test_isEven()
```
]]

---
# Write the test function _first_!

.leftcol[.code80[
### Step 1: Write the test function
```{r}
test_isEven <- function() {
    cat("Testing isEven()...")
    stopifnot(isEven(1) == FALSE)
    stopifnot(isEven(2) == TRUE)
    stopifnot(isEven(-7) == FALSE)
    cat("Passed!\n")
}
```
]]
--
.rightcol[.code80[
### Step 2: Write the function
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
### Step 3: Test the function
```{r}
test_isEven()
```
]]

---
# Test cases to consider

.leftcol40[
- ### Normal cases
- ### Large & small cases
- ### Pairs of opposites
- ### Edge cases
- ### Special cases
]

---
# Test cases to consider

.leftcol40[
- ### **.red[Normal cases]**
- ### Large & small cases
- ### Pairs of opposites
- ### Edge cases
- ### Special cases
]
.rightcol60[.code80[
### Example:
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
]
- `isEven(1) == FALSE`
- `isEven(2) == TRUE`
- `isEven(-7) == FALSE`
]

---
# Test cases to consider

.leftcol40[
- ### Normal cases
- ### **.red[Large & small cases]**
- ### Pairs of opposites
- ### Edge cases
- ### Special cases
]
.rightcol60[.code80[
### Example:
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
]
- `isEven(8675309) == FALSE`
- `isEven(-8675309) == FALSE`
]

---
# Test cases to consider

.leftcol40[
- ### Normal cases
- ### Large & small cases
- ### **.red[Pairs of opposites]**
- ### Edge cases
- ### Special cases
]
.rightcol60[.code80[
### Example:
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
]
Need cases that return both `TRUE` and `FALSE`
- `isEven(52) == TRUE`
- `isEven(53) == FALSE`
]

---
# Test cases to consider

.leftcol40[
- ### Normal cases
- ### Large & small cases
- ### Pairs of opposites
- ### **.red[Edge cases]**
- ### Special cases
]
.rightcol60[.code80[
### Example:
```{r}
isPositive <- function(n) {
    return(n > 0)
}
```
]
- `isPositive(0.000001) == TRUE`
- `isPositive(0) == FALSE`
- `isPositive(-0.000001) == FALSE`
]

---
# Test cases to consider

.leftcol40[
- ### Normal cases
- ### Large & small cases
- ### Pairs of opposites
- ### Edge cases
- ### **.red[Special cases]**
]
.rightcol60[
- Negative numbers
- `0` and `1` for integers
- The empty string, `""`
- Strange input _types_, e.g. `"2"` instead of `2`.
]

---
# Testing function inputs

--
.leftcol45[.code80[
What if we gave `isEven()` the wrong input type?
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
```{r error=TRUE}
isEven('42')
```
]]
--
.rightcol55[.code80[
An improved function that checks inputs:
```{r}
isEven <- function(n) {
    if (! is.numeric(n)) {
        return(NaN)
    }
    return((n %% 2) == 0)
}
```
]]

---
# Testing function inputs

.leftcol45[.code80[
What if we gave `isEven()` the wrong input type?
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
```{r error=TRUE}
isEven('42')
```
]]
.rightcol55[.code80[
An improved function that checks inputs:
```{r}
isEven <- function(n) {
    if (! is.numeric(n)) {
        return(NaN)
    }
    return((n %% 2) == 0)
}
```
```{r}
isEven('42')
isEven(TRUE)
```
]]

---
class: inverse

```{r, echo=FALSE}
countdown(minutes = 15,
          warn_when = 30,
          update_every = 1,
          top = 0,
          right = 0,
          font_size = '2em')
```

# Think-Pair-Share

For each of the following functions, start by writing a test function that tests the function for a variety of values of inputs. Consider cases that you might not expect!

.leftcol[
1. Write the function `isFactor(f, n)` that takes two integer values and returns `TRUE` if `f` is a factor of `n`, and `FALSE` otherwise. Note that every integer is a factor of `0`. Assume `f` and `n` will only be numeric values, e.g. `2` is a factor of `6`.
]
.rightcol[
2. Write the function `isMultiple(m, n)` that takes two integer values and returns `TRUE` if `m` is a multiple of `n` and `FALSE` otherwise. Note that `0` is a multiple of every integer other than itself. Hint: You may want to use the `isFactor(f, n)` function you just wrote above. Assume `m` and `n` will only be numeric values.
]
