
class: inverse, middle

# .fancy[Final Exam Review]

## 1. .orange[R basics]
## 2. Data wrangling
## 3. Data visualization

---

## Operators: Relational (`=`, `<`, `>`, `<=`, `>=`) and Logical (`&`, `|`, `!`)

```{r}
x <- FALSE
y <- FALSE
z <- TRUE
```

a Write a logical statement that compares the objects `x`, `y`, and `z` and returns `TRUE`

b) Fill in **relational** operators to make this statement return `TRUE`:

`! (x __ y) & ! (z __ y)`

c) Fill in **logical** operators to make this statement return `FALSE`:

`! (x __ y) | (z __ y)`

---
# Numeric Data

--
.leftcol[
Doubles:
```{r}
typeof(3.14)
```
]
--
.rightcol[
"Integers":
```{r}
typeof(3)
```
]

---

# Actual Integers

```{r}
typeof(3L)
```
--
Check if a number is an "integer":
```{r}
n <- 3
is.integer(n) # Doesn't work!
```
--
```{r}
n == as.integer(n) # Compare n to a converted version of itself
```

---

# Logical Data

`TRUE` or `FALSE`

--
```{r}
x <- 1
y <- 2
```

--
```{r}
x > y # Is x greater than y?
```

--
```{r}
x == y
```

---

## Tricky data type stuff

--
.leftcol[
Logicals become numbers when doing math
```{r}
TRUE + 1 # TRUE becomes 1
```
```{r}
FALSE + 1 # FALSE becomes 0
```
```{r}
sum(c(TRUE, FALSE, TRUE))
```

]
--
.rightcol[
Be careful of accidental strings
```{r}
typeof("3.14")
typeof("TRUE")
```
]

---

# Basic function syntax

.code90[
```{r eval=FALSE}
functionName <- function(arguments) {
    # Do stuff here
    return(something)
}
```
]

---

# Basic function syntax

In English:

> "`functionName` is a `function` of `arguments` that does..."

.code90[
```{r eval=FALSE}
functionName <- function(arguments) {
    # Do stuff here
    return(something)
}
```
]

---

# Basic function syntax

Example:

> "`squareRoot` is a `function` of `n` that...returns the square root of `n`"

```{r}
squareRoot <- function(n) {
    return(n^0.5)
}
```

--

```{r}
squareRoot(64)
```

---

# Test function syntax

--
.leftcol[
### Function:
```{r, eval=FALSE}
functionName <- function(arguments) {
    # Do stuff here
    return(something)
}
```
]
--
.rightcol[
### Test function:
```{r, eval=FALSE}
test_functionName <- function() {
    cat("Testing functionName()...")
    # Put test cases here
    cat("Passed!\n")
}
```
]

---

# Writing test cases with `stopifnot()`

`stopifnot()` stops the function if whatever is inside the `()` is not `TRUE`.

--
.leftcol[
### Function:
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}
```
- `isEven(1)` should be `FALSE`
- `isEven(2)` should be `TRUE`
- `isEven(-7)` should be `FALSE`
]
--
.rightcol[
### Test function:
```{r}
test_isEven <- function() {
    cat("Testing isEven()...")
    stopifnot(isEven(1) == FALSE)
    stopifnot(isEven(2) == TRUE)
    stopifnot(isEven(-7) == FALSE)
    cat("Passed!\n")
}
```
]

---

# When testing _numbers_, use `almostEqual()`

```{r}
addTwo <- function(n1, n2) {
    return(n1 + n2)
}
```

--
.leftcol[
A **bad** test function
```{r, error=TRUE}
test_addTwo <- function() {
    cat("Testing addTwo()...")
    stopifnot(addTwo(0.1, 0.2) == 0.3) #<<
    cat("Passed!\n")
}

test_addTwo()
```
]
--
.rightcol[
A **good** test function
```{r, include=FALSE}
almostEqual <- function(n1, n2, threshold = 0.00001) {
    return(abs(n1 - n2) <= threshold)
}
```
```{r}
test_addTwo <- function() {
    cat("Testing addTwo()...")
    stopifnot(almostEqual( #<<
      addTwo(0.1, 0.2), 0.3)) #<<
    cat("Passed!\n")
}

test_addTwo()
```
]

---

.leftcol[
### Use `for` loops when the number of iterations is **_known_**.

1. Build the sequence
2. Iterate over it
```{r}
for (i in 1:5) { # Define the sequence #<<
    cat(i, '\n')
}
```
]
--
.rightcol[
### Use `while` loops when the number of iterations is **_unknown_**.

1. Define stopping condition
2. Manually increase condition
```{r}
i <- 1
while (i <= 5) { # Define stopping condition #<<
    cat(i, '\n')
    i <- i + 1 # Increase condition #<<
}
```
]

---

# The universal vector generator: `c()`

--
.cols3[
## Numeric vectors
```{r}
x <- c(1, 2, 3)
x
```
]
--
.cols3[
## Character vectors
```{r}
y <- c('one', 'two', 'three')
y
```
]
--
.cols3[
## Logical vectors
```{r}
z <- c(TRUE, FALSE, TRUE)
z
```
]

---

# Elements in vectors must be the same type

### Type hierarchy:
- `character` > `numeric` > `logical`
- `double` > `integer`


--
.cols3[
Coverts to characters:
```{r}
c(1, "foo", TRUE)
```
]
--
.cols3[
Coverts to numbers:
```{r}
c(7, TRUE, FALSE)
```
]
--
.cols3[
Coverts to double:
```{r}
c(1L, 2, pi)
```
]

---

# Most functions operate on vector _elements_

--
```{r}
x <- c(3.14, 7, 10, 15)
round(x)
```
--
Works on custom functions too:
```{r}
isEven <- function(n) {
    return((n %% 2) == 0)
}

isEven(x)
```

---

# Comparing vectors

Check if 2 vectors are the same:

```{r}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
```
```{r, eval=FALSE}
x == y
```
--
```{r, echo=FALSE}
x == y
```

---

# Comparing vectors with `all()` and `any()`

--
.leftcol[
`all()`: Check if _all_ elements are the same
```{r}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
all(x == y)
```
```{r}
x <- c(1, 2, 3)
y <- c(-1, 2, 3)
all(x == y)
```
]
--
.rightcol[
`any()`: Check if _any_ elements are the same
```{r}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
any(x == y)
```
```{r}
x <- c(1, 2, 3)
y <- c(-1, 2, 3)
any(x == y)
```
]
