
class: inverse, middle

# .fancy[Final Exam Review]

## 1. R basics
## 2. Data wrangling
## 3. .orange[Data visualization]

---

# Making plot layers with ggplot2

<br>
### 1. The data
### 2. The aesthetic mapping (what goes on the axes?)
### 3. The geometries (points? bars? etc.)

---

# Layer 1: The data

The `ggplot()` function initializes the plot with whatever data you're using

.leftcol[
```{r ggblank, fig.show='hide'}
ggplot(data = bears)
```
]
.rightcol[.blackborder[
```{r ref.label='ggblank', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---

# Layer 2: The aesthetic mapping

The `aes()` function determines which variables will be _mapped_ to the geometries<br>(e.g. the axes)

.leftcol[
```{r ggaes, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age))
```
]
.rightcol[.blackborder[
```{r ref.label='ggaes', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---

# Layer 3: The geometries

Use `+` to add geometries (e.g. points)

.leftcol[
```{r ggpoint, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() #<<
```
]
.rightcol[.blackborder[
```{r ref.label='ggpoint', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---

# Layer 3: The geometries

Can also map aesthetics within the geom

.leftcol[
```{r ggpoint_aes, fig.show='hide'}
ggplot(data = bears) +
    geom_point(aes(x = year, y = age)) #<<
```
]
.rightcol[.blackborder[
```{r ref.label='ggpoint_aes', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---

## Change the point color

.leftcol[
```{r ggbearsblue, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point(color = 'blue') #<<
```
]
.rightcol[.blackborder[
```{r ref.label='ggbearsblue', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---

## Change the point color<br>**based on another variable**:

.leftcol[
```{r ggbearsage, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point(aes(color = gender)) #<<
```
]
.rightcol[.blackborder[
```{r ref.label='ggbearsage', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---

## Make bar charts with `geom_col()`

Piping directly into ggplot

.leftcol[
```{r bears_geom_col_pipe, fig.show='hide'}
bears %>%
  count(month) %>%  # Pipe into ggplot #<<
  ggplot() +
  geom_col(aes(x = month, y = n))
```
]
.rightcol[
```{r ref.label='bears_geom_col_pipe', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---

## If you can't read the axes...flip them

.leftcol[
```{r birds_geom_col, fig.show='hide'}
birds %>%
  mutate(phase =
    str_to_lower(phase_of_flt)) %>%
  count(phase) %>%
  ggplot() +
  geom_col(aes(x = phase, y = n)) #<<
```
]
.rightcol[
```{r ref.label='birds_geom_col', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---

## If you can't read the axes...flip them

.leftcol[
```{r birds_geom_col_flip, fig.show='hide'}
birds %>%
  mutate(phase =
    str_to_lower(phase_of_flt)) %>%
  count(phase) %>%
  ggplot() +
  geom_col(aes(y = phase, x = n)) #<<
```
]
.rightcol[
```{r ref.label='birds_geom_col_flip', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---

## Use `reorder()` to sort an axis

.leftcol[
```{r birds_geom_col_sort, fig.show='hide'}
birds %>%
  mutate(phase =
    str_to_lower(phase_of_flt)) %>%
  count(phase) %>%
  ggplot() +
  geom_col(aes(y = reorder(phase, n), x = n)) #<<
```
]
.rightcol[
```{r ref.label='birds_geom_col_sort', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---

### Change bar width:  `width`
### Change bar color: `fill`
### Change bar outline: `color`

.leftcol[
```{r bears_geom_col_colors, fig.show='hide'}
bears %>%
  count(month) %>%
  ggplot() +
  geom_col(aes(x = month, y = n),
           width = 0.7, #<<
           fill = "blue", #<<
           color = "red") #<<
```
]
.rightcol[
```{r ref.label='bears_geom_col_colors', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---

### Map the `fill` to `bearType`

.leftcol[.code70[
```{r bears_geom_col_fill, fig.show='hide'}
bears %>%
  count(month, bearType) %>% #<<
  ggplot() +
  geom_col(aes(x = month, y = n,
               fill = bearType)) #<<
```
]

Note that I had to summarize the count by both `month` and `bearType`

.code70[
```{r, eval=FALSE}
bears %>%
  count(month, bearType)
```
]
.code60[
```{r, echo=FALSE}
bears %>%
  count(month, bearType) %>%
  arrange(month, bearType)
```
]]
.rightcol[
```{r ref.label='bears_geom_col_fill', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]

---

# "Factors" = Categorical variables

By default, R makes numeric variables _continuous_

.leftcol[
```{r, eval=FALSE}
bears %>%
  count(month) %>%
  ggplot() +
  geom_col(aes(x = month, y = n))
```
**The variable `month` is a _number_**
]
.rightcol[.blackborder[
```{r ref.label='bears_geom_col_pipe', echo=FALSE, fig.height=5, fig.width=7, fig.align='center'}
```
]]

---

# "Factors" = Categorical variables

You can make a continuous variable _categorical_ using `as.factor()`

.leftcol[.code80[
```{r bears_geom_col_factor, fig.show='hide'}
bears %>%
  count(month) %>%
  ggplot() +
  geom_col(aes(x = as.factor(month), #<<
               y = n))
```
]

**The variable `month` is a _factor_**
]
.rightcol[.blackborder[
```{r ref.label='bears_geom_col_factor', echo=FALSE, fig.height=5, fig.width=7}
```
]]

---
# Working with themes

Themes change _global_ features of your plot, like the background color, grid lines, etc.

--
.leftcol[
```{r ggbears_basic, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point()
```
]
.rightcol[
```{r ref.label='ggbears_basic', echo=FALSE, fig.height=5, fig.width=7}
```
]

---

# Working with themes

Themes change _global_ features of your plot, like the background color, grid lines, etc.

.leftcol[
```{r ggbears_theme, fig.show='hide'}
ggplot(data = bears,
       aes(x = year, y = age)) +
    geom_point() +
    theme_bw() #<<
```
]
.rightcol[
```{r ref.label='ggbears_theme', echo=FALSE, fig.height=5, fig.width=7}
```
]
---

# Save ggplots with `ggsave()`

--
First, assign the plot to an object name:
```{r, eval=FALSE}
scatterPlot <- ggplot(data = bears) +
    geom_point(aes(x = year, y = age))
```

--
Then use `ggsave()` to save the plot:
```{r, eval=FALSE}
ggsave(filename = here('plots', 'scatterPlot.png'),
       plot   = scatterPlot,
       width  = 6, # inches
       height = 4)
```
